Workday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population

### 2023-12-11 manuscript/paper.Rmd
hi

\begin{table}[H]
\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients based on Model 5. These results are adjusted for other significant potentially confounding factors based on univariate regression analysis.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
\midrule
\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
\textbf{Sleep} & 0.75 & (0.50, 1.00) & 0.002 \\
$<6$hrs &  &  &  \\
$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
\textbf{Age} & -0.04& (-0.22, 0.13) & 0.6 \\
\textbf{TC} & -3.5  & (-5.2, -1.8) & <0.001 \\
\textbf{TC*Age} & 0.11  & (0.07, 0.14) & <0.001 \\
\bottomrule
\end{tabular}
\end{table}
```

```{r tabfour, echo=FALSE, eval=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

@@

The current study investigated the
intricate interplay between age, sleep, cholestrol levels and systolic blood
pressure adjusting for other potentially confounding covariates informed by literature. We performed comprehensive model comparisons to explore the association between age, total cholestrol levels, sleep durations and systolic blood pressure. We find based on the best model, that work day sleep durations ranging between 6 and 8 hours and more than 8 hours are associated with a decrease in systolic blood pressure relative to the 6 hour reference group. Additionally, our study reports substantial effect modification of the effect of total cholestrol levels on systolic blood pressure by age whe adjusted for selected confounders within the study population. Previous studies exploring sleep and systolic blood pressure have found similar associations [@su2022association].

>>>>

\begin{table}[H]
\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients based on Model 5. These results are adjusted for other significant potentially confounding factors based on univariate regression analysis.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
\midrule
\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
\textbf{Sleep} & 0.75 & (0.50, 1.00) & 0.002 \\
$<6$hrs &  &  &  \\
$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
\textbf{Age} & -0.04& (-0.22, 0.13) & 0.6 \\
\textbf{TC} & -3.5  & (-5.2, -1.8) & <0.001 \\
\textbf{TC*Age} & 0.11  & (0.07, 0.14) & <0.001 \\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabfour, echo=FALSE, eval=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

@@

<!-- The current study investigated the intricate interplay between age, sleep, cholestrol levels and systolic blood pressure adjusting for other potentially confounding covariates informed by literature. We performed comprehensive model comparisons to explore the association between age, total cholestrol levels, sleep durations and systolic blood pressure. We find based on the best model, that work day sleep durations ranging between 6 and 8 hours and more than 8 hours are associated with a decrease in systolic blood pressure relative to the 6 hour reference group. Additionally, our study reports substantial effect modification of the effect of total cholestrol levels on systolic blood pressure by age whe adjusted for selected confounders within the study population. Previous studies exploring sleep and systolic blood pressure have found similar associations [@su2022association]. -->

The present investigation explote into the settle interplay among age,
sleep patterns, cholesterol levels, and systolic blood pressure, while
accounting for various potentially confounding factors gleaned from
pertinent literature. Our study conducted an exhaustive series of model
comparisons to elucidate the intricate relationship between age, total
cholesterol levels, sleep durations, and systolic blood pressure.

In light of our findings derived from the best model based on the
performance measures in Table \@ref(tab:tabmodel), we observed that
weekday sleep durations spanning between 6 and 8 hours, as well as
durations exceeding 8 hours, exhibit a decrease in systolic blood
pressure in comparison to the reference group of individuals sleeping
for 6 hours. Furthermore, our investigation revealed a noteworthy effect
modification regarding the influence of total cholesterol levels on
systolic blood pressure when adjusted for specific confounding variables
within the studied cohort, especially in relation to age.

Our study contributes to the existing body of knowledge by
substantiating the associations uncovered in previous research
concerning sleep patterns and systolic blood pressure
[@su2022association]. This correlation underscores the significance of
sleep parameters as influential factors in understanding blood pressure
dynamics.
